---
title: "Practical Machine Learning Course Project"
author: "Petr Minin"
date: '22 םגאנ 2018 ד '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```

## Executive Summary
The goal of the Practical Machine Learning Course Paper is to predict one of the 5 different ways of performing barbell lifts by 6 different participants by using data from accelerometers on the belt, forearm, arm, and dumbell. 
The Random Forest Model shows the 100% accuracy both on the training and the test set. It correctly predicts the way of performing barbell lifts in all cases.

## Downloading and Proccesing Data
###R Code
```{r}
library(caret)
library(randomForest)
set.seed(123)

#download data
training <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
control <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

#split the training set in two
inTrain = createDataPartition(training$classe, p = 3/4)[[1]]
training = training[inTrain,]
testing = training[-inTrain,]

#deleting variables, that are not taken from acceeleromter
training=training[,-c(1:7)]

#deleting variables with >50% NA
na_count <-sapply(training, function(x) sum(is.na(x)))
na_count <- na_count<0.5*nrow(training)
training <- training[,na_count]

#deleting variables with>50% empty values
empty_count <- sapply(training, function(x) sum(x ==""))
empty_count <- empty_count<0.5*nrow(training)
training <- training[,empty_count]
```

###Downloading and Proccessing Data
The training dataset was downloaded from the url: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The control dataset was downloaded from the url:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The training dataset was split into a training and test dataset in proportion 3:1.

The following preprocessing steps were undertaken with the training dataset:
1. The variables that were not taken from the accelerometer were deleted;
2. The variables with more than 50% of NAs were deleted;
3. The variables with more than 50% of the missing values were deleted.

The final training dataset consists of 14719 observations of 53 variables, one of which is the class variable to be predicted.

##Fitting a Random Forest Model
###R Code
```{r}
#fitting random forest model
RF_Model <- randomForest(classe~., data=training)

#Insample and out of sample error
predict_training <- predict(RF_Model, newdata=training)
predict_testing <- predict(RF_Model, newdata=testing)

#Predicting for the control set
predict_control <- predict(RF_Model, newdata=control)

```

###Fitting a Random Forest Model

The randomForest library was used to fit a random forest model. The fitted model showes a 100% accuracy both - on the training dataset:

``` {r , echo=FALSE}
confusionMatrix(training$classe, predict_training)
```

and on the testing dataset:

``` {r , echo=FALSE}
confusionMatrix(testing$classe, predict_testing)
```
The in-sample error and out-of-sample error is 0.

Finally we use the model to predict the class of performing barbell lifts for the control dataset:

``` {r , echo=FALSE}
predict_control
```
